---
title: " What Impacts do Artificial Oyster Reef have on Land Erosion Caused by Storm Surges in the Gulf Coast? " 
Author: "Lydia Clark and Rachel Hendricks"
date: "2023-03-30"
output: github_document

---
```{r}
library(tidyverse)
library(readr)
library(skimr)
library(dplyr)

```

# Import Data Manually 

**Don't need hurricane due to being already completed**

*Storm Surge Dataset* i
```{r read files}
stormsurge <- read_csv("globalpeaksurgedb.csv")
Encrusting_organism <- read_csv("2017-2019 Encrusting organism count and density (1).csv")
Oyster_shell <- read_csv("2017-2019 Oyster shell heights (1).csv")
```



**SKIM DATA**

```{r}
skim(Oyster_shell)
skim(Encrusting_organism)
```

```{r}
skim(stormsurge)
```


**DATA ANALYSIS**

```{r fileter for US} 
stormsurge %>%
filter(Country == "US")
```

```{r filtering for specific states}
US_stormsurge <-stormsurge %>%
  filter(Country== "US")

Gulf_stormsurge <- stormsurge %>%
  filter(State == "AL" | State =="TX"| State == "LA" | State =="MS" | State == "FL")


Gulf_stormsurge

```

Bar chart comparing storm surge heights along states. We can see that Florida has more frequent surges at higher heights than any other state. Florida in general experiences more surges than the other states, while Mississippi and Alabama experience the least. Could use this in damage assessments. 

```{r surge}

ggplot(Gulf_stormsurge, aes(y = Surge_m, fill = State)) +
  geom_bar()+
    facet_wrap(~ State, nrow = 1, labeller = label_wrap_gen(width = 12)) +
  labs(
    y = "Storm Surge Height (m)",
    title = "Comparision of Storm Surge Heights in States along Gulf Coast") + 
  theme(legend.position = "none") 

```


ggplot(Gulf_stormsurge, aes(y = Storm_Tide_m, fill = State)) +

 
Bar plot working in progress. Trying to figure out how to make the chart more readable. 

```{r tide}
ggplot(Gulf_stormsurge, aes(x = Storm_Tide_m, fill = State)) +
  geom_bar()+
    facet_wrap(~ State, nrow = 1, labeller = label_wrap_gen(width = 12)) +
  labs(
    y = "Storm Tide Height (m) ",
    title = "Comparision of Storm Tide Heights in States along Gulf Coast") + 
  theme(legend.position = "none") 
```

Box plot showing state vs storm tide. Same idea as the bar chart below, but shown differently. 
```{r box plot}
ggplot(Gulf_stormsurge, aes(x = State, y = Storm_Tide_m)) +
  geom_boxplot()+
  labs(
    x = "State", y = "Storm Tide Height (m)", 
    title = "Comparision of Storm Tide Heights in States along Gulf Coast") + 
  theme(legend.position = "none") 
```


This graph shows that Florida experiences higher storm tides than other states, and Alabama has the lowest. Could be useful when connecting damage caused by storm to occurrence. 

```{r tide height and state }
ggplot(Gulf_stormsurge, aes(x = State, fill= Surge_ft)) +
  geom_bar()+
  labs(
    x = "State Where Storm Occurred ", y = "Surge Height", 
    title = "Comparision of Storm Tide Heights in States along Gulf Coast") 
```

Trying to look at relationship between oyster density and site. Need to understand what the site locations are so we can connect that to specific states. 

```{r}
ggplot(Encrusting_organism, aes(x = SITE, y = OYSTER_DENSITY, fill = OYSTER_COUNT)) +
  geom_boxplot()+
  labs(
    x = "Site ", y = "Oyster Density", 
    title = "Number of oysters in Site") +
  theme(legend.position = "none") 
```

#Mapping# 
````{r global_options, include = FALSE}
knitr::opts_chunk$set(
    message = FALSE,
    warning = FALSE,
    cache = TRUE,
    tidy = FALSE,
    fig.width = 7,
    fig.height = 7,
    out.width = "65%")

library(tidyverse)
library(sf)
library(maps)
library(tools)
library(ggspatial)

```

```{r}
LA_stormsurge <- Gulf_stormsurge %>%
  filter(State == ("LA"))
```

```{r eval= FALSE}
stormsurge_sites <- tibble(longitude = c(-93.7001,-90.0506,-89.1708,-90.5914,-89.9017,-91.2081,-90.8349
,-89.9873,-92.6301,-90.4231,-93.3300,-89.8672,-90.0689,-89.6786,-89.5297,-93.4525,-92.1075,-91.3833,-89.5189,-93.3442,-90.6712,-89.7565,-90.0292,-91.8772,-90.6633
,-90.6537,-91.6683,-93.5625,-89.8576,-92.1469,-89.8576,-89.8576,-90.6693,-89.9873,-89.7274, -91.4206,-89.8576,-90.4268,-93.3433,-93.1837,-89.6060,-89.6183,-90.0370,-89.7659) +
(latitude = c(29.7632,29.2097,29.0169,29.2542,29.5797,29.6886,29.2554,29.2633,29.5953,30.1075,29.7400,30.1275,30.3528,29.8551,29.3683,29.7697,29.7451,29.3667,29.3825,29.7622,29.2357,29.5450,30.0157,29.7161,29.2453,29.1627,29.7452,29.7608,30.0038,30.0038,29.5875,30.0038,30.0038,29.2305,29.2366,30.1733,29.5742,30.0038,30.1076,29.7683,29.7867,29.5858,29.5983,30.0310,29.5666)) +

stormsurge_sites <- st_as_sf(LA_sites, coords = c("longitude", "latitude"), crs = 4326) +

states <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE)) %>%
  mutate(ID = toTitleCase(ID)) +

counties <- st_as_sf(maps::map("county", plot = FALSE, fill = TRUE)) %>%
  filter(str_detect(ID, "Louisiana")) %>%
  mutate(area = as.numeric(st_area(st_make_valid(geom)))) +

parish <- tibble(state = rep("Louisiana", 8), 
                 city = c("Cameron", "Vermilion", "Iberia", "St Mary", "Terrebonne", "Lafourche", "Plaquemines", "St Bernard"), 
                 latitude = c(29.797721,29.860141,	29.499263,29.502976, 29.229968, 29.69523,	29.324005,29.879791 ), 
                 longitude = c( -93.325154,-92.381362,-91.753882,	-91.443469,-90.753281,-90.525782,-89.474218,-89.322695)) +

parish <- st_as_sf(parish, coords = c("longitude", "latitude"), crs = 4326) +


ggplot() +
#Add the states polygons and make the fill color light yellow-ish (I use hex codes to define colors)
  geom_sf(data = states, 
          fill = "#f4f1de") +
#add the counties polygons, make the fill color light orange and the line color light gray
  geom_sf(data = counties, 
          fill = "#f2cc8f", 
          color = "#787878") +
#add the sites points, make the point size 4, shape #23 (diamond), and the fill color dark orange
  geom_sf(data = stormsurge_sites, 
          size = 4, shape = 23, 
          fill = "#e07a5f") +
#add the cities points, make the point size 2 and color a medium gray
  geom_sf(data = parish, 
          size = 2, 
          color = "#383838") +
#add state names using `geom_sf_text()`; make the font size 4, the text color medium gray, and the fontface bold
  geom_sf_text(data = states, aes(label = ID), 
               size = 4, 
               color = "#383838", 
               fontface = "bold") +
#add the site name labels using `geom_sf_label_repel()`; "nudge" the labels a bit so they don't look too squished together
  geom_sf_label_repel(data = stormsurge_sites, aes(label = name),
                      nudge_x = c(0.5, 1), 
                      nudge_y = c(1, 0)) +
#add city labels using `geom_sf_text_repel()`; make the font color medium gray, font size 3, fontface italic, and nudge the labels a bit so they don't overlap with the counties polygons as much
  geom_sf_text_repel(data = cities, aes(label = city), 
                     color = "#383838", 
                     size = 3, 
                     fontface = "italic",
                     nudge_x = c(0.5, 1, 1, -0.5),
                     nudge_y = c(0, 0.1, 0.1, 0)) +
#add a scale bar; color the lines and bars dark gray and fill the bar with bluish-green
  annotation_scale(location = "bl", 
                   width_hint = 0.4, 
                   bar_cols = c("#383838", "#d8e2dc"), 
                   line_col = "#383838", 
                   text_col = "#383838") +
#add a north-facing arrow; fill the arrow with dark gray and bluish-green and color the lines dark gray
  annotation_north_arrow(location = "bl", 
                         which_north = "true", 
                         pad_y = unit(0.25, "in"), 
                         style = north_arrow_fancy_orienteering(fill = c("#d8e2dc", "#383838"),
                                                                line_col = "#383838", 
                                                                text_col = "#383838")) +
  coord_sf(xlim = c(-88.032338,-94.443539), 
           ylim = c(28.264494,33.310300), 
           expand = FALSE) +
  labs(x = "Longitude", 
       y = "Latitude", 
       title = "Study Sites", 
       subtitle = "Palm Beach County, Florida") +
  theme(panel.background = element_rect(fill = "#d8e2dc"))


```


